Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow

### 2012-12-17 Text/Final/introduction.tex
Minor corrections in the text (mostly the new additions, but also some old parts).

Typical numerical studies of mass transfer
\cite{kreutzer-overview,vanbaten-circular} do not consider the simulation of
bubble shapes for $Ca>0.1$. {\color{red} Interesting work by \citet{onea-mass} only takes into the account short lengths bubble in the squared microchannels. As well the Schmidt number around $0.8$ is  used and the thickness of the concentration and momentum boundary layer is similar leading to qualitative results.} The usual simulation of mass transfer is performed as
follows: 
\begin{description}
 \item[I] The bubble shape is calculated either
through analytical correlations \cite{bretherton} or experimental correlations
\cite{cerro-bubble-train} without directly resolving bubble shapes through
multiphase simulations.  The expressions for bubble shapes are available only
for flows with capillary number $Ca<0.1$.
  \item[II] Hydrodynamic fields
are then obtained by performing simulations of one-component flow around the
bubble by imposing the bubble velocity on the channel walls. Thus, the simulations are
performed in the reference frame moving with the bubble. A stress-free
condition is imposed at the bubble surface.
  \item[III]  The mass transfer
simulations are performed in the reference frame moving with the bubble. The
saturation concentration is imposed at the bubble surface. Only one unit cell
containing a single bubble is used for simulations. Periodic concentration boundary conditions are
utilized to determine the volumetric mass transfer coefficient, which is
calculated through the following equation \cite{vanbaten-circular}:
\begin{equation} \label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\cinoutlet\rangle}
\frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\ cell\ volume}},
\end{equation}
 where $\langle\cinoutlet(t)\rangle=\int{C \uinoutlet
\mathrm{d}A}/\int{\uinoutlet\mathrm{d}A}$ is the space-averaged inlet/outlet (periodic boundary conditions)
concentration as a function of time. Therefore, in terms of the mass transfer definition, $\langle\cinoutlet(t)\rangle$ plays the role of the 
characteristic concentration.
%This is the characteristic concentration in the definition, Eq. \ref{eq:main:definition}.  
The time-averaged
concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference
between the overall average concentration in the whole domain
($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$) at time $t_1$ and at time
$t_2$ divided by the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and experimental correlations of \citet{bercic-mass}
% TODO: was good where?
was good. 
 \end{description}

@@

While it is clear that periodic boundary conditions can be employed for the calculation
 of hydrodynamic fields, the same does not apply to the mass transfer coefficient
simulations. Experimental correlations \cite{bercic-mass} show that the
concentration in a bubble train along the streamwise direction changes exponentially with distance.
Mass transfer simulations however, are made only for one unit cell
using periodic boundary conditions with the same concentration at the inlet
and at the outlet. The question is how single unit cell simulation corresponds to
experimental measurements arises where concentration difference is measured at
the distances of at least a few unit cells \cite{bercic-mass}. In
other words, one needs to understand how the discrete one unit cell simulation
corresponds to the continuous picture in experiments where one does not
distinguish discrete bubbles but takes measurements of concentration at different
locations.

@@

To establish numerical procedures we performed multiphase simulations
to extract bubble shapes
\cite{kuzmin-binary2d,kuzmin-binary3d} for the range of capillary numbers
$Ca=0.1\div 1.0$. For this range of capillary numbers
we were able to capture the bubble shape change and the change of hydrodynamic
patterns. The mass transfer simulations presented here were performed with various boundary
conditions (open, periodic) and with a few unit cells ($1$ to $10$ unit cells).
As our numerical approach we take the lattice Boltzmann method, a relatively
new CFD competitor developed during last $20$ years
\cite{frisch,mcnamara,HJ,HSB}. This method was successfully applied to simulate
not only single phase hydrodynamic problems \cite{yu}, but also multiphase
flows \cite{Shan-chen:extended,swift,gunstensen}, heat transfer
\cite{yuan-thermal,zhang-thermal}, and ferrofluids \cite{dellar-ferro,kuzmin-aniso}.

@@

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
to the bubble train flow to derive expressions to connect the space- and time-averages.  Then, the lattice
Boltzmann model used to simulate mass transfer is presented, followed by benchmarks. Finally,
numerical simulations of various boundary conditions and simulations spanning a few unit cells for
different hydrodynamic patterns are presented to establish the procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also presented.

@@

\subsection{Bubble train}
In the application to the bubble train flow it is useful to think of one bubble as a point
source to be able to use the calculations presented above. For example, the expression
(\ref{main:mass:transfer:expression}) was used in experiments by
\citet{bercic-mass}. However, one should be accurate with the definition of velocities because two
different phases co-exist in the bubble train flow. Usually, one can take the velocity $U$ to be
a bulk velocity or $U=\ugas+\uliq$, where $\ugas$ and $\uliq$ are liquid and gas
superficial velocities, respectively. {\color{red} Gas and liquid superficial and actual velocities are connected with each other through complicated dependency on $Ca$, $\epsilon$ \cite{kuzmin-binary3d}.}

@@

In comparison with analytical calculations and simplifications, the numerical approach can take into
account the complex mixing patterns and geometries. However, there are challenges as  how to
mimic the continuous picture where the medium is moving with bulk velocity $U=\ugas+\uliq$  as it is done in
experiments. Thus, the questions indicated in Section \ref{intro} arise.  The next section gives more
 details about numerical simulations.

@@

\subsection{TRT $D2Q9$ model}
The lattice Boltzmann equation (LBE) operates on a square/cubic grid representing the
physical domain {\color{red} in the non-dimensional space} (hereafter we use subscript LB to denote non-dimensional quantities). It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another along  
directions specified by a discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)= f_i(\bm{x},t)-\omega \bigl(f_i(\bm{x},t)-eq_i(\bm{x},t)\bigr),&&\text{
collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
 $eq_i$ is the equilibrium probability distribution function, and $\omega$ is the
relaxation parameter. The term $-\omega \bigl(f_i- eq_i\bigr)$ is the so-called BGK
collision operator \cite{bgk}. However, the approach used here is the TRT
(two-relaxation-times) collision operator \cite{ginzburg-main,ginzburg-saturated-flow}. In
comparison with the widely used BGK collision operator, the TRT collision operator has better accuracy
for diffusion and convection fluxes, as well as a larger range of parameters where the scheme is stable.

@@

The equilibrium functions for the $D2Q9$ TRT model are represented as \cite{kuzmin-stability-optimal}:
\begin{equation}
\begin{aligned}
&eq_i^{+}=eq_i^{(m)}+g^{(u)} eq_i^{(u)}\\
&eq_i^{(m)}=t_i^{(m)} c_e+ eq_i^{(a)}\\
&eq_i^{(u)}=t_i^{(u)} \frac{u_{LB,x}^2+u_{LB,y}^2}{2}+\frac{u_{LB,x}^2-u_{LB,y}^2}{4} p_i^{(xx)}+g_{xy}^{(u)}\frac{u_{LB,x}
u_{LB,y}}{4} p_i^{xy}\\
&eq_i^{(a)}=\frac{K_{LB,xx}-K_{LB,yy}}{4} p_i^{xx}+\frac{K_{LB,xy}}{4} p_i^{(xy)}\\
&eq_i^{-}=t_i^{(a)}  u_{LB,\alpha} c_{i\alpha},
\end{aligned}
\end{equation}
where $K_{LB,xx,yy,xy}$ are proportional to components of the diffusion tensor,
$c_e=\frac{K_{LB,xx}+K_{LB,yy}}{2}$, parameters $g^{(u)}$ and $g^{u}_{xy}$ are either zero or one (see
below), the tensor $p_i^{(xx)}=c_{ix}^2-c_{iy}^2$, the tensor $p_i^{(xy)}=c_{ix} c_{iy}$, the
weights
$t_i^{(u,m,a)}$ can be chosen based on stability criteria. The most commonly used set of weights, the so-called
``hydrodynamic`` weights, were chosen:
\begin{equation}
t_i^{(u)}=t_i^{(m)}=t_i^{(a)}=\Bigl\{0,\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{12},
\frac {1}{12},\frac{1}{12},\frac{1}{12}\Bigr\}
\end{equation}

@@

\subsection{Poiseulle velocity profile}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,\pm \delta)=\cstar,\, \frac{\partial C}{\partial y }(x,0)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq
The procedure to solve this problem is presented in Appendix \ref{appendix:poiseuille} which yields
the final solution as:
\begin{equation}
C=\cstar-\cstar \sum_{m=0}{C_m e^{-m^4 \frac{x}{\delta}\frac{1}{Pe}} e^{-m^2
y^2/(2\delta^2)}{_1F_1}\Bigl(-\frac{m^2}{4}+\frac{1}{4},\frac{1}{2},m^2 \frac{y^2}{\delta^2}\Bigr)},
\end{equation}
where coefficients $C_m$ are taken from Eq. \ref{coeff:series:parabolic:profile}. The comparison
between contours of analytical and simulation results is presented in Fig.
\ref{fig:parabolic:comparison}. Parameters were taken as: $D=0.0185$, the grid dimension is $80\times1600$. The centerline velocity is  $U_0=0.05$
which yields the Peclet number $Pe=U_0 \delta / D= 108.108$.
The results are in good agreement. The simulations capture accurately the singular derivative for $x=0$.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{parabolic_profile_comparison.eps}
\caption{Comparison between the analytical concentration contours and simulations with pressure
anti bounce-back conditions, Eq. \ref{antibb}. The simulation was done for $D_{LB}=0.0185$ with a
$80\times1600$ grid. The centerline velocity is $U_0=0.05$, and the Peclet number is $108.108$. \label{fig:parabolic:comparison}}
\end{figure}

Now the LBM is validated against the benchmarks relevant for the flow around bubbles, one can
examine the cases mentioned in Section
\ref{section:cases} to calculate the volumetric mass transfer coefficient for Taylor bubble train flow.

\section{Numerical approach}
\label{sec:numerics}
A multiphase code was utilized to obtain the flow patterns and bubble shapes for different capillary numbers
\cite{kuzmin-binary2d}. Five particular cases were chosen to be examined, their results are summarized
in Table \ref{table:capillary:cases}. 
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$\ububble$ &$\delta$&$\holdup$
&$\uliq$&$\ugas$&$\lbubble$&$\lslug$\\
\hline
$0.097$ &$1.656$  &$0.0055$ &$0.092$ &$0.30$ &$0.0046$&$0.0016$&$5.79$&$9.21$\\ 
$0.254$ &$4.318$  &$0.0143$ &$0.132$ &$0.28$ &$0.0108$&$0.0041$&$6.12$&$8.88$\\ 
$0.526$ &$8.938$  &$0.0297$ &$0.157$ &$0.27$ &$0.0209$&$0.0080$&$6.19$&$8.81$\\
$0.750$ &$12.744$ &$0.0424$ &$0.167$ &$0.25$ &$0.0293$&$0.0107$&$5.96$&$9.04$\\
$1.040$ &$17.665$ &$0.0588$ &$0.177$ &$0.22$ &$0.0397$&$0.0135$&$5.59$&$9.41$\\
\hline
\end{tabularx}
\caption{Sample results with the binary liquid lattice Boltzmann model \cite{kuzmin-binary2d}. The
following notations are used: the capillary number $Ca=\frac{\ububble L}{\rho \nu_{\mathrm{liq}}}$, $\uliq$ is the superficial liquid velocity, $\ugas$ is the
superficial gas velocity, $\holdup$ is the gas holdup. $\delta$ is the
non-dimensional film thickness, $\lbubble$ and $\lslug$ are the non-dimensional bubble and slug lengths (defined in channel heights).The simulation sketch is presented in Fig.
\ref{fig:benchmark:hydro}. \label{table:capillary:cases}}
\end{table}
Note that the velocities (LB system) in Table \ref{table:capillary:cases} are small. This means that to
match large Peclet numbers, $Pe=\frac{\ububble L}{D}$, usually used in experiments, one needs to decrease the diffusion coefficient
$D=\frac{1}{3}\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}\Bigr)$. Thus, the parameter $\omegaminus\approx 0.5$. However, for such values of 
$\omegaminus$ the stability of the lattice Boltzmann method drastically decreases
\cite{kuzmin-d1q3}. On the other hand, one iteration in the lattice Boltzmann system corresponds
to a physical time step  $\Delta t_{\mathrm{phys}}=U_{\mathrm{bubble,LB}} \frac{\Delta
x}{U_{\mathrm{bubble,phys}}}$, {\color{red} where $U_{\mathrm{bubble,phys}}$ is the physical velocity $\bigl[\mathrm{m^2}/\mathrm{s}\bigr]$}, and $t_{\mathrm{phys}}$ is the physical time step $\bigl[\mathrm{s}\bigr]$. The iteration time is proportional to the velocity $U_{\mathrm{LB}}$
and the typical number of simulation steps to obtain the steady-state mass transfer coefficient for
$Ca<0.2$ is of the order of a few million. Therefore, it is desirable to increase    $U_{\mathrm{LB}}$
while maintaining the Peclet number. If one increases the velocity, then
$\omegaminus$ increases as well, which impacts positively on the stability of the LBM.

Given all the considerations above, mass transfer simulations are performed as follows:
\begin{description}
 \item[Flow field] Given a capillary number $Ca$, one needs to obtain hydrodynamic fields around
the bubble using the multiphase binary liquid lattice Boltzmann model according to our previous work
\cite{kuzmin-binary2d}. Periodic boundary conditions were used in that work. The grid used  was
$202\times 3000$ which corresponds to the fluid domain  of size $200\times3000$. That grid resolution was
taken to ensure grid
independency of the results \cite{kuzmin-binary2d}. Note that we do not approximate bubble shapes by correlations, but we directly resolve them by using the multiphase solver. 
 \item[Bubble reference frame] Once the hydrodynamics is solved, the mass transfer simulations
are conducted in the reference frame moving with the bubble, where the bubble stands still and the liquid
flows around the
bubble. We impose a uniform and steady concentration on the surface
of the bubble with the anti bounce-back condition, Eq. \ref{antibb}.
 \item[Velocity improvement] One can scale the velocity to
perform faster simulations. However, before doing it one needs to improve the velocity field.
This issue arises because of the
multiphase model used in the flow simulations. The binary liquid lattice Boltzmann model is a diffuse
interface model where no clear boundary between gas and liquid exists.
% TODO: Is this only a problem of how we determine the bubble shape or also a problem of
% spurious velocities within the model itself?
We obtain the bubble shape by imposing a condition on the order parameter field $\phi$ with $\phi\leq0$ in the bubble \cite{kuzmin-binary2d}. The velocity of the
bubble is defined as the bubble tip velocity. Because of the square grid, the shape of the bubble is determined within an accuracy of one grid spacing. Thus,
there is an error in the determination of the bubble velocity. Though these errors are small,
there is still a small non-zero velocity component pointing into the bubble in some places, see Fig. 8 in
\cite{kuzmin-binary2d} where some streamlines are penetrating the bubble surface.
This small velocity is amplified upon the velocity scaling and is inconsistent with the
advection-diffusion equation leading to instability after many iterations.

>>>>

Typical numerical studies of mass transfer
\cite{kreutzer-overview,vanbaten-circular} do not consider the simulation of
bubble shapes for $Ca>0.1$. {\color{red} The interesting work by
\citet{onea-mass} only takes into account short length bubbles in microchannels
with a square cross section. As well the Schmidt number around $0.8$ is  used
and the thickness of the concentration and momentum boundary layer is similar
leading to qualitative results.} The usual simulation of mass transfer is
performed as follows: 
\begin{description}
 \item[I] The bubble shape is calculated either
through analytical correlations \cite{bretherton} or experimental correlations
\cite{cerro-bubble-train} without directly resolving bubble shapes through
multiphase simulations.  The expressions for bubble shapes are available only
for flows with capillary number $Ca<0.1$.
  \item[II] Hydrodynamic fields
are then obtained by performing simulations of one-component flow around the
bubble by imposing the bubble velocity on the channel walls. Thus, the simulations are
performed in the reference frame moving with the bubble. A stress-free
condition is imposed at the bubble surface.
  \item[III]  The mass transfer
simulations are performed in the reference frame moving with the bubble. The
saturation concentration is imposed at the bubble surface. Only one unit cell
containing a single bubble is used for simulations. Periodic concentration boundary conditions are
utilized to determine the volumetric mass transfer coefficient, which is
calculated through the following equation \cite{vanbaten-circular}:
\begin{equation} \label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\cinoutlet\rangle}
\frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\ cell\ volume}},
\end{equation}
 where $\langle\cinoutlet(t)\rangle=\int{C \uinoutlet
\mathrm{d}A}/\int{\uinoutlet\mathrm{d}A}$ is the space-averaged inlet/outlet (periodic boundary conditions)
concentration as a function of time. Therefore, in terms of the mass transfer definition, $\langle\cinoutlet(t)\rangle$ plays the role of the 
characteristic concentration.
%This is the characteristic concentration in the definition, Eq. \ref{eq:main:definition}.  
The time-averaged
concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference
between the overall average concentration in the whole domain
($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$) at time $t_1$ and at time
$t_2$ divided by the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and experimental correlations of \citet{bercic-mass}
% TODO: was good where?
was good. 
 \end{description}

@@

While it is clear that periodic boundary conditions can be employed for the calculation
 of hydrodynamic fields, the same does not apply to the mass transfer coefficient
simulations. Experimental correlations \cite{bercic-mass} show that the
concentration in a bubble train along the streamwise direction changes exponentially with distance.
Mass transfer simulations however, are made only for one unit cell
using periodic boundary conditions with the same concentration at the inlet
and at the outlet. The question how a single unit cell simulation corresponds to
experimental measurements arises where the concentration difference is measured at
distances of at least a few unit cells \cite{bercic-mass}. In
other words, one needs to understand how the discrete one unit cell simulation
corresponds to the continuous picture in experiments where one does not
distinguish discrete bubbles but takes measurements of concentration at different
locations.

@@

To establish numerical procedures we performed multiphase simulations
to extract bubble shapes
\cite{kuzmin-binary2d,kuzmin-binary3d} for the range of capillary numbers
$Ca=0.1\div 1.0$. For this range of capillary numbers
we were able to capture the bubble shape change and the change of hydrodynamic
patterns. The mass transfer simulations presented here were performed with various boundary
conditions (open, periodic) and with a few unit cells ($1$ to $10$ unit cells).
As our numerical approach we take the lattice Boltzmann method, a relatively
new CFD competitor developed during the last $20$ years
\cite{frisch,mcnamara,HJ,HSB}. This method was successfully applied to simulate
not only single phase hydrodynamic problems \cite{yu}, but also multiphase
flows \cite{Shan-chen:extended,swift,gunstensen}, heat transfer
\cite{yuan-thermal,zhang-thermal}, and ferrofluids \cite{dellar-ferro,kuzmin-aniso}.

@@

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
to the bubble train flow to derive expressions to connect the space- and time-averages.  Then, the lattice
Boltzmann model used to simulate mass transfer is presented, followed by benchmarks. Finally,
numerical simulations of various boundary conditions and simulations spanning a few unit cells for
different hydrodynamic patterns are presented to establish the procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also discussed.

@@

\subsection{Bubble train}
In the application to the bubble train flow it is useful to think of one bubble as a point
source to be able to use the calculations presented above. For example, the expression
(\ref{main:mass:transfer:expression}) was used in experiments by
\citet{bercic-mass}. However, one should be accurate with the definition of velocities because two
different phases co-exist in the bubble train flow. Usually, one can take the velocity $U$ to be
a bulk velocity or $U=\ugas+\uliq$, where $\ugas$ and $\uliq$ are liquid and gas
superficial velocities, respectively. {\color{red} The gas and liquid superficial and actual velocities are connected with each other through a complex dependency on $Ca$, $\epsilon$ \cite{kuzmin-binary3d}.}

@@

In comparison with analytical calculations and simplifications, the numerical approach can take into
account the complex mixing patterns and geometries. However, there are challenges as to how to
mimic the continuous picture where the medium is moving with bulk velocity $U=\ugas+\uliq$  as it is done in
experiments. Thus, the questions indicated in Section \ref{intro} arise.  The next section gives more
 details about numerical simulations.

@@

\subsection{TRT $D2Q9$ model}
The lattice Boltzmann equation (LBE) operates on a square/cubic grid representing the
physical domain {\color{red} in the non-dimensional space} (hereafter we use subscript LB to denote non-dimensional quantities). It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. The LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another along  
directions specified by a discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)= f_i(\bm{x},t)-\omega \bigl(f_i(\bm{x},t)-eq_i(\bm{x},t)\bigr),&&\text{
collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
 $eq_i$ is the equilibrium probability distribution function, and $\omega$ is the
relaxation parameter. The term $-\omega \bigl(f_i- eq_i\bigr)$ is the so-called BGK
collision operator \cite{bgk}. However, the approach used here is the TRT
(two-relaxation-times) collision operator \cite{ginzburg-main,ginzburg-saturated-flow}. In
comparison with the widely used BGK collision operator, the TRT collision operator has better accuracy
for diffusion and convection fluxes, as well as a larger range of parameters where the scheme is stable.

@@

The equilibrium functions for the $D2Q9$ TRT model are represented as \cite{kuzmin-stability-optimal}:
\begin{equation}
\begin{aligned}
&eq_i^{+}=eq_i^{(m)}+g^{(u)} eq_i^{(u)}\\
&eq_i^{(m)}=t_i^{(m)} c_e+ eq_i^{(a)}\\
&eq_i^{(u)}=t_i^{(u)} \frac{u_{LB,x}^2+u_{LB,y}^2}{2}+\frac{u_{LB,x}^2-u_{LB,y}^2}{4} p_i^{(xx)}+g_{xy}^{(u)}\frac{u_{LB,x}
u_{LB,y}}{4} p_i^{xy}\\
&eq_i^{(a)}=\frac{K_{LB,xx}-K_{LB,yy}}{4} p_i^{xx}+\frac{K_{LB,xy}}{4} p_i^{(xy)}\\
&eq_i^{-}=t_i^{(a)}  u_{LB,\alpha} c_{i\alpha},
\end{aligned}
\end{equation}
where $K_{LB,xx,yy,xy}$ are proportional to components of the diffusion tensor,
$c_e=\frac{K_{LB,xx}+K_{LB,yy}}{2}$, parameters $g^{(u)}$ and $g^{u}_{xy}$ are either zero or one (see
below), and the tensor $p_i^{(xx)}=c_{ix}^2-c_{iy}^2$, the tensor $p_i^{(xy)}=c_{ix} c_{iy}$, the
weights
$t_i^{(u,m,a)}$ can be chosen based on stability criteria. The most commonly used set of weights, the so-called
``hydrodynamic`` weights, were chosen:
\begin{equation}
t_i^{(u)}=t_i^{(m)}=t_i^{(a)}=\Bigl\{0,\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{12},
\frac {1}{12},\frac{1}{12},\frac{1}{12}\Bigr\}
\end{equation}

@@

\subsection{Poiseuille velocity profile}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,\pm \delta)=\cstar,\, \frac{\partial C}{\partial y }(x,0)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq
The procedure to solve this problem is presented in Appendix \ref{appendix:poiseuille} which yields
the final solution as:
\begin{equation}
C=\cstar-\cstar \sum_{m=0}{C_m e^{-m^4 \frac{x}{\delta}\frac{1}{Pe}} e^{-m^2
y^2/(2\delta^2)}{_1F_1}\Bigl(-\frac{m^2}{4}+\frac{1}{4},\frac{1}{2},m^2 \frac{y^2}{\delta^2}\Bigr)},
\end{equation}
where coefficients $C_m$ are taken from Eq. \ref{coeff:series:parabolic:profile}. The comparison
between contours of analytical and simulation results is presented in Fig.
\ref{fig:parabolic:comparison}. Parameters were taken as: $D=0.0185$, the grid dimension is $80\times1600$. The centerline velocity is  $U_0=0.05$
which yields the Peclet number $Pe=U_0 \delta / D= 108.108$.
The results are in good agreement. The simulations capture accurately the singular derivative for $x=0$.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{parabolic_profile_comparison.eps}
\caption{Comparison between the analytical concentration contours and simulations with pressure
anti bounce-back conditions, Eq. \ref{antibb}. The simulation was done for $D_{LB}=0.0185$ with a
$80\times1600$ grid. The centerline velocity is $U_0=0.05$, and the Peclet number is $108.108$. \label{fig:parabolic:comparison}}
\end{figure}

Now that the LBM is validated against the benchmarks relevant for the flow around bubbles, one can
examine the cases mentioned in Section
\ref{section:cases} to calculate the volumetric mass transfer coefficient for the Taylor bubble train flow.

\section{Numerical approach}
\label{sec:numerics}
A multiphase code was utilized to obtain the flow patterns and bubble shapes for different capillary numbers
\cite{kuzmin-binary2d}. Five particular cases were chosen to be examined, their results are summarized
in Table \ref{table:capillary:cases}. 
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$\ububble$ &$\delta$&$\holdup$
&$\uliq$&$\ugas$&$\lbubble$&$\lslug$\\
\hline
$0.097$ &$1.656$  &$0.0055$ &$0.092$ &$0.30$ &$0.0046$&$0.0016$&$5.79$&$9.21$\\ 
$0.254$ &$4.318$  &$0.0143$ &$0.132$ &$0.28$ &$0.0108$&$0.0041$&$6.12$&$8.88$\\ 
$0.526$ &$8.938$  &$0.0297$ &$0.157$ &$0.27$ &$0.0209$&$0.0080$&$6.19$&$8.81$\\
$0.750$ &$12.744$ &$0.0424$ &$0.167$ &$0.25$ &$0.0293$&$0.0107$&$5.96$&$9.04$\\
$1.040$ &$17.665$ &$0.0588$ &$0.177$ &$0.22$ &$0.0397$&$0.0135$&$5.59$&$9.41$\\
\hline
\end{tabularx}
\caption{Sample results with the binary liquid lattice Boltzmann model \cite{kuzmin-binary2d}. The
following notations are used: the capillary number $Ca=\frac{\ububble L}{\rho \nu_{\mathrm{liq}}}$, $\uliq$ is the superficial liquid velocity, $\ugas$ is the
superficial gas velocity, $\holdup$ is the gas holdup. $\delta$ is the
non-dimensional film thickness, $\lbubble$ and $\lslug$ are the non-dimensional bubble and slug lengths (defined as multiplies of the channel height). The simulation sketch is presented in Fig.
\ref{fig:benchmark:hydro}. \label{table:capillary:cases}}
\end{table}
Note that the velocities (LB system) in Table \ref{table:capillary:cases} are small. This means that to
match large Peclet numbers, $Pe=\frac{\ububble L}{D}$, usually used in experiments, one needs to decrease the diffusion coefficient
$D=\frac{1}{3}\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}\Bigr)$. Thus, the parameter $\omegaminus\approx 0.5$. However, for such values of 
$\omegaminus$ the stability of the lattice Boltzmann method drastically decreases
\cite{kuzmin-d1q3}. On the other hand, one iteration in the lattice Boltzmann system corresponds
to a physical time step  $\Delta t_{\mathrm{phys}}=U_{\mathrm{bubble,LB}} \frac{\Delta
x}{U_{\mathrm{bubble,phys}}}$, {\color{red} where $U_{\mathrm{bubble,phys}}$ is the physical velocity $\bigl[\mathrm{m}/\mathrm{s}\bigr]$}, and $t_{\mathrm{phys}}$ is the physical time step $\bigl[\mathrm{s}\bigr]$. The iteration time is proportional to the velocity $U_{\mathrm{LB}}$
and the typical number of simulation steps to obtain the steady-state mass transfer coefficient for
$Ca<0.2$ is of the order of a few million. Therefore, it is desirable to increase    $U_{\mathrm{LB}}$
while maintaining the Peclet number. If one increases the velocity, then
$\omegaminus$ increases as well, which impacts positively on the stability of the LBM.

Given all the considerations above, mass transfer simulations are performed as follows:
\begin{description}
 \item[Flow field] Given a capillary number $Ca$, one needs to obtain hydrodynamic fields around
the bubble using the multiphase binary liquid lattice Boltzmann model according to our previous work
\cite{kuzmin-binary2d}. Periodic boundary conditions were used in that work. The grid used  was
$202\times 3000$ which corresponds to the fluid domain  of size $200\times3000$. That grid resolution was
taken to ensure grid
independency of the results \cite{kuzmin-binary2d}. Note that we do not approximate bubble shapes by correlations, but directly resolve them using the multiphase solver. 
 \item[Bubble reference frame] Once the hydrodynamics is solved, the mass transfer simulations
are conducted in the reference frame moving with the bubble, where the bubble stands still and the liquid
flows around the
bubble. We impose a uniform and steady concentration on the surface
of the bubble with the anti bounce-back condition, Eq. \ref{antibb}.
 \item[Velocity improvement] One can scale the velocity to
perform faster simulations. However, before doing it one needs to improve the velocity field.
This issue arises because of the
multiphase model used in the flow simulations. The binary liquid lattice Boltzmann model is a diffuse
interface model where no clear boundary between gas and liquid exists.
% TODO: Is this only a problem of how we determine the bubble shape or also a problem of
% spurious velocities within the model itself?
We obtain the bubble shape by imposing a condition on the order parameter field $\phi$ with $\phi\leq0$ in the bubble \cite{kuzmin-binary2d}. The velocity of the
bubble is defined as the bubble tip velocity. Because of the square grid, the shape of the bubble is determined within an accuracy of one grid spacing. Thus,
there is an error in the determination of the bubble velocity. Though these errors are small,
there is still a small non-zero velocity component pointing into the bubble in some places, see Fig. 8 in
\cite{kuzmin-binary2d} where some streamlines are penetrating the bubble surface.
This small velocity is amplified upon the velocity scaling and is inconsistent with the
advection-diffusion equation leading to instability after many iterations.

Thus, before performing the mass transfer simulations an additional single phase hydrodynamic
simulation is performed. A free-surface solver was developed in order to obtain a velocity field
consistent with the advection-diffusion equation. We take results from the
multiphase simulations, extract a
bubble shape using the phase indicator $\phi\leq0$, and approximate the bubble shape by the stair-case
line with imposed free-slip boundary condition on it. The obtained bubble velocity is then imposed on the walls. This corresponds to conducting 
simulations in the reference frame moving with the bubble. Appendix \ref{appendix-free-surface} covers the simple
free-slip boundary condition implementation drastically improving velocity patterns. The system is
iterated until a
steady state is reached
and as a result, all non-zero velocity components perpendicular to the bubble surface are completely eliminated.
Note, that these types of simulations are much faster than the original
multiphase simulations.  We compared original multiphase simulations with
one-component free-slip simulations. All quantities such as superficial slug and liquid velocities are
within $3\%$ for the capillary number in the range $0.05\leq Ca \leq 1.0$. 
One can see in Fig. \ref{fig:streamlines:tweaked:velocity} two streamline profiles for $Ca=0.097$  and $Ca=1.040$.
\begin{figure}[htb!]
\includegraphics[angle=90,width=\textwidth]{performed_ca0097_out.eps}\\
\includegraphics[angle=90,width=\textwidth]{performed_ca1040_out.eps}
\caption{The streamline patterns produced by the free-surface flow solver with simplified approximation
of the free-slip bubble surface, see Appendix \ref{appendix-free-surface}. Two completely different
velocity patterns are obtained, $Ca=0.097$ (top) and $Ca=1.040$ (bottom).
\label{fig:streamlines:tweaked:velocity}}
\end{figure}
